---
title: "Time Series Analysis & Forecasting"
author: "18BCE1104 - Ankita Duraphe"
date: "11/02/2021"
output:
  word_document: default
  pdf_document: default
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

### Forecasting on sample data

* * *
highest Petrol price in every month from 2018 to yesterday

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
data_vector = c(75.63, 76.12, 76.29, 77.47, 81.47, 81.32, 79.8, 81.63, 86.8, 87.39, 82.51, 75.34,74.03, 74.43, 75.65, 75.96, 75.9, 74.39, 76.29, 75.65, 77.34, 77.56, 77.763, 78.1, 79.04, 76.01, 74.79, 72.37, 75.78, 83.63, 83.67, 85.05, 85.09, 84.19, 85.31, 86.51, 88.8, 89.68)
```

* * * 
converting vector to dataset

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
data_forecasting = ts(data_vector, start=c(2018,1), end=c(2021,2), frequency=12)
```

* * * 
importance of splitting automatically

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
start(data_forecasting)
end(data_forecasting)
frequency(data_forecasting)
cycle(data_forecasting)
summary(data_forecasting)
```

* * * 
Analysis PART: plotting the series

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
plot(data_forecasting)
```

* * * 
regression

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
plot(data_forecasting)
abline(reg=lm(data_forecasting~time(data_forecasting)), col="blue")
```

* * * 
month-wise (range and mean) plot 

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
monthplot(data_forecasting)
```

* * * 
year-wise mean plot

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
plot(aggregate(data_forecasting,FUN=mean))
```

* * * 
month-wise box plot

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
boxplot(data_forecasting~cycle(data_forecasting))
```

* * *
Load the required packages

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
library(fabletools)
library(ggplot2)
library(forecast)
library(ggfortify)
```

* * * 
visualisation

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
decompose_fc = decompose(data_forecasting,"multiplicative")
autoplot(decompose_fc)
```

* * * 
FORECASTING / PREDICTING

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
library(forecast)
```

* * * 
Monthwise predicting

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
seasonplot(data_forecasting)
```

* * * 
Autocorrelation and Partial Autocorrelation plots

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
acf(data_forecasting)   #for MA - Q
pacf(data_forecasting, lag=length(data_forecasting),pl=TRUE) # for AR - P
```

* * * 
ARIMA

AR (P - autoregressive lags) + I (d - order of differentiation) + MA (Q - moving average) \
correlation between previous time period with current \
noise | averaging / smoothing

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
fit_arima = arima(data_forecasting, order=c(3, 2, 2))
accuracy(fit_arima)        
newdata_arima = forecast(fit_arima, 4)
plot(newdata_arima)
```

* * * 
**MFE: Mean Forecast Error or ME: Mean Error (for forecast)**

From accuracy(fit_arima), MFE or ME = 0.01807052 

* * * 
**MAD: Mean Absolute Deviation or MAE: Mean Absolute Error**

From accuracy(fit_arima), MAD or MAE = 1.632346 

* * * 
Auto ARIMA 

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
fit_autoarima = auto.arima(data_forecasting)
accuracy(fit_autoarima)
newdata_autoarima = forecast(fit_autoarima, 4)
plot(newdata_autoarima)
```

* * * 
**MFE: Mean Forecast Error or ME: Mean Error (for forecast)**

From accuracy(fit_autoarima), MFE or ME = 0.3113201  

* * * 
**MAD: Mean Absolute Deviation or MAE: Mean Absolute Error**

From accuracy(fit_autoarima), MAD or MAE = 1.679884 

### Forecasting for Weather History dataset

* * * 
Load the required packages

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
library(ggplot2)
library(dplyr)
```

* * * 
Read the dataset

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
data <- read.csv("weatherHistory.csv")
```

* * * 
View the dataset

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
View(data)
```

* * * 
Converting To Time Series object using ts() function: For Weekly data

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
weather_forecasting = ts(data$Humidity, start=c(2006,1,4), end=c(2016,9,9), frequency=7)
```

* * * 
importance of spliting automatically

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
start(weather_forecasting)
end(weather_forecasting)
frequency(weather_forecasting)
cycle(weather_forecasting)
summary(weather_forecasting)
```

* * * 
Analysis PART: plotting the series

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
plot(weather_forecasting)
```

* * * 
regression

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
plot(weather_forecasting)
abline(reg=lm(weather_forecasting~time(weather_forecasting)), col="blue")
```

* * * 
month-wise (range and mean) plot 

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
monthplot(weather_forecasting)
```

* * * 
year-wise mean plot

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
plot(aggregate(weather_forecasting,FUN=mean))
```

* * * 
month-wise box plot

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
boxplot(weather_forecasting~cycle(weather_forecasting))
```

* * * 
visualisation

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
decompose_fc = decompose(weather_forecasting,"multiplicative")
autoplot(decompose_fc)
```

* * * 
FORECASTING / PREDICTING

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
library(forecast)
```

* * * 
Monthwise predicting

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
seasonplot(weather_forecasting)
```

* * * 
Autocorrelation and Partial Autocorrelation plots

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
acf(weather_forecasting)   #for MA - Q
pacf(weather_forecasting, lag=length(weather_forecasting),pl=TRUE) # for AR - P
```

* * * 
ARIMA

AR (P - autoregressive lags) + I (d - order of differentiation) + MA (Q - moving average) \
correlation between previous time period with current \
noise | averaging / smoothing

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
fit_arima = arima(weather_forecasting, order=c(3, 2, 2))
accuracy(fit_arima)        
newdata_arima = forecast(fit_arima, 4)
plot(newdata_arima)
```

* * * 
**MFE: Mean Forecast Error or ME: Mean Error (for forecast)**

From accuracy(fit_arima), MFE or ME = 0.00494041   

* * * 
**MAD: Mean Absolute Deviation or MAE: Mean Absolute Error**

From accuracy(fit_arima), MAD or MAE = 0.04667016  

* * * 
Auto ARIMA 

```{r, warning=FALSE, message=FALSE, tidy=TRUE}
fit_autoarima = auto.arima(weather_forecasting)
accuracy(fit_autoarima)
newdata_autoarima = forecast(fit_autoarima, 4)
plot(newdata_autoarima)
```

* * * 
**MFE: Mean Forecast Error or ME: Mean Error (for forecast)**

From accuracy(fit_autoarima), MFE or ME = 0.0005175949    

* * * 
**MAD: Mean Absolute Deviation or MAE: Mean Absolute Error**

From accuracy(fit_autoarima), MAD or MAE = 0.04887203 

* * *
**Conclusion:** \
Forecasting, including MFE & MAD, has been successfully performed on sample data as well as on weatherHistory dataset.